ABC Group Level Fits

### 2024-01-22 Model/abc_hier.qmd
abc indv fit parameterized reports

walk(c("Display_Functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
ids <- c(1,2,4,5,6,7,8, 10,11,12,13)
ids2 <- c(1,66,141,142,36,57,81)
ids3 <- c(20,71,101,4,76,192)

@@

ind_fits <- here::here(readRDS("data/ind_abc_ss_1000_ng100_bufp55"))
ind_fits <- ind_fits |> list_assign(runInfo = zap())

# names(ind_fits)
# names(ind_fits$id_fits_alm_varied)
# names(ind_fits$id_fits_alm_varied[[1]])

teter_results_df <- imap_dfr(ind_fits, ~imap_dfr(.x, ~.x[["teter_results"]] ,.id="id"))

combined_results_df <- imap_dfr(ind_fits, ~imap_dfr(.x, ~{
  teter_data <- .x[["teter_results"]] %>% mutate(result_type = "teter_results")
  te_data <- .x[["te_results"]] %>% mutate(result_type = "te_results")
  tr_data <- .x[["tr_results"]] %>% mutate(result_type = "tr_results")
  combined_data <- bind_rows(teter_data, te_data, tr_data)
}, .id = "id"))

id1 <- combined_results_df |> filter(id %in% ids2) 
names(id1)
typeof(id1$sim_dat)
str(id1$sim_dat[[1]])
# sim_dat is a list dataframe. Add the value of rank into sim_dat

id1 <- id1 |> select(rank,sim_dat) |> unnest(sim_dat) |>filter(rank<=2)

@@

```{r, fig.width=7, fig.height=6}

plt_c_gi <- combined_results_df |> 
  ggplot(aes(x=c)) + geom_density(aes(fill=Group), alpha=.5) + 
  facet_wrap(Fit_Method~Model,scales="free",ncol=2) + 
  theme_bw() + theme(legend.position = "bottom") + labs(title = "Posterior distribution of c")

plt_lr_gi <- combined_results_df |> 
  ggplot(aes(x=lr)) + geom_density(aes(fill=Group), alpha=.5) + 
  facet_wrap(Fit_Method~Model,scales="free",ncol=2) + 
  theme_bw() + theme(legend.position = "bottom") + labs(title = "Posterior distribution of lr")

plt_c_gi + plt_lr_gi

@@

```{r fig.width=11, fig.height=18}

post_dat_long <- combined_results_df |> filter(id %in% (c(ids2,ids3)), rank<=1) |> 
  select(sim_dat) |> unnest(sim_dat) |> filter(expMode2=="Test") |>
  mutate(facet_label = paste0("Model: ", Model, "\n", "Fit: ", Fit_Method, "\n", "c: ", round(c, 4), "\n", "lr: ", round(lr, 4))) |>
  pivot_longer(cols = c(y,pred), names_to = "Resp", values_to = "val")

post_dat <- combined_results_df |> filter(id %in% (c(ids2,ids3)), rank<=5) |> 
  select(sim_dat) |> unnest(sim_dat) |> filter(expMode2=="Test")

post_dat_long |> filter(id %in% ids2) |> ggplot(aes(x = x, y = val, fill=as.factor(Model))) + 
  stat_summary(fun=mean, geom="bar", position=position_dodge(), alpha=.75) +
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge()) +
  #stat_halfeye(position=position_dodge()) +
  facet_wrap(~id, scales="free") + labs(title = "Average of posterior predictive distributions")

@@

>>>>

pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, stringr)
conflict_prefer_all("dplyr", quiet = TRUE)

walk(c("Display_Functions"), ~ source(here::here(paste0("Functions/", .x, ".R"))))
ds <- readRDS(here::here("data/e1_md_11-06-23.rds"))  |> as.data.table()
ids <- c(1,2,4,5,6,7,8, 10,11,12,13)
ids2 <- c(1,66,36)
ids3 <- c(20,71,101,4,76,192)
idsBad <- c(76,192, 101)

@@

ind_fits <- readRDS(here::here("data/indv_sim/ind_abc_ss_1000_ng100_bufp55.rds"))

#ind_fits <- ind_fits |> list_assign(runInfo = zap())
# names(ind_fits)
# names(ind_fits$id_fits_alm_varied)
# names(ind_fits$id_fits_alm_varied[[1]])

#teter_results_df <- imap_dfr(ind_fits, ~imap_dfr(.x, ~.x[["teter_results"]] ,.id="id"))

ind_fits_df <- imap_dfr(ind_fits |> list_assign(runInfo = zap()), ~imap_dfr(.x, ~{
  teter_data <- .x[["teter_results"]] %>% mutate(result_type = "teter_results")
  te_data <- .x[["te_results"]] %>% mutate(result_type = "te_results")
  tr_data <- .x[["tr_results"]] %>% mutate(result_type = "tr_results")
  combined_data <- bind_rows(teter_data, te_data, tr_data)
}, .id = "id"))

post_dat <- ind_fits_df |> 
  select(sim_dat,rank) |> unnest(sim_dat) |> filter(expMode2=="Test") |>
  group_by(id,condit,x) |>
  mutate(avg_y=mean(y)) |> 
  relocate(id,condit,Model,Fit_Method,expMode2,tr,rank,c,lr,x,y,avg_y)

post_dat_l <- post_dat %>%
  pivot_longer(
    cols = c(pred, y),
    names_to = "Resp",
    values_to = "val"
  ) %>%
  mutate(
    Resp = case_when(
      Resp == "y" ~ "Observed",
      Model == "ALM" ~ "ALM",
      Model == "EXAM" ~ "EXAM"
    )
  ) |> arrange(tr,id,Resp)

id1 <- ind_fits_df |> filter(id %in% ids2) 
names(id1)
typeof(id1$sim_dat)
str(id1$sim_dat[[1]])
id1 <- id1 |> select(rank,sim_dat) |> unnest(sim_dat) |>filter(rank<=2)

@@

##  Posterior Average
```{r fig.width=12, fig.height=17}

et <- post_dat_l |>
  group_by(id, condit, Fit_Method, Resp, x) |>
  summarise(val = mean(val), .groups = 'drop') |>
  pivot_wider(
    names_from = Resp,
    values_from = val,
    values_fill = list(val = NA)
  ) |>
  mutate(
    ALM_error = abs(ALM - Observed),
    EXAM_error = abs(EXAM - Observed),
    Best_Model = case_when(
      ALM_error < EXAM_error ~ "ALM",
      EXAM_error < ALM_error ~ "EXAM",
      TRUE ~ NA_character_  # In case of a tie or missing data
    )
  )

et_counts <- et %>%
  group_by(condit,Fit_Method, Best_Model) %>% 
  summarise(count = n(), .groups = 'drop')

 ggplot(et, aes(x = factor(x), fill = Best_Model)) +
  geom_bar(position = "dodge") +
  facet_wrap(condit~Fit_Method, scales = "free") +
  labs(
    title = "Frequency of Best Model by x Value",
    x = "x Value",
    y = "Frequency",
    fill = "Best Model"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(et, aes( x = factor(x), y = id, fill = Best_Model)) +
  geom_tile(color = "white") +
  facet_wrap(condit~Fit_Method,scales="free") +
  labs(title = "Best Model for Each ID and X Value",
       x = "X",
       y = "ID",
       fill = "Best Model") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

et_sum_x <- post_dat_l |>
  group_by(condit, Fit_Method, Resp, x) |>
  summarise(val = mean(val), .groups = 'drop') |>
  pivot_wider(
    names_from = Resp,
    values_from = val,
    values_fill = list(val = NA)
  ) |>
  group_by(condit, Fit_Method, x) |>
  transmute(
    ALM_error = abs(ALM - Observed),
    EXAM_error = abs(EXAM - Observed),
    Best_Model = case_when(
      ALM_error < EXAM_error ~ "ALM",
      EXAM_error < ALM_error ~ "EXAM",
      TRUE ~ NA_character_  # In case of a tie or missing data
    )
  )

et_sum_id <- post_dat_l |>
  group_by(id,condit, Fit_Method, Resp) |>
  summarise(val = mean(val), .groups = 'drop') |>
  pivot_wider(
    names_from = Resp,
    values_from = val,
    values_fill = list(val = NA)
  ) |>
  group_by(id,condit, Fit_Method) |>
  transmute(
    ALM_error = abs(ALM - Observed),
    EXAM_error = abs(EXAM - Observed),
    Best_Model = case_when(
      ALM_error < EXAM_error ~ "ALM",
      EXAM_error < ALM_error ~ "EXAM",
      TRUE ~ NA_character_  # In case of a tie or missing data
    )
  )

et_sum <- post_dat_l |>
  group_by(id,condit, Fit_Method, Resp) |>
  summarise(val = mean(val), .groups = 'drop') |>
  pivot_wider(
    names_from = Resp,
    values_from = val,
    values_fill = list(val = NA)
  ) |>
  mutate(
    ALM_error = abs(ALM - Observed),
    EXAM_error = abs(EXAM - Observed),
    Best_Model = case_when(
      ALM_error < EXAM_error ~ "ALM",
      EXAM_error < ALM_error ~ "EXAM",
      TRUE ~ NA_character_  # In case of a tie or missing data
    )
  ) |>
  group_by(condit, Fit_Method) %>%
  summarise(
    Avg_ALM_error = mean(ALM_error, na.rm = TRUE),
    Avg_EXAM_error = mean(EXAM_error, na.rm = TRUE),
    N_ALM = sum(Best_Model == "ALM", na.rm = TRUE),
    N_EXAM = sum(Best_Model == "EXAM", na.rm = TRUE)
  ) %>%
  mutate(
    Best_Model = case_when(
      Avg_ALM_error < Avg_EXAM_error ~ "ALM",
      Avg_EXAM_error < Avg_ALM_error ~ "EXAM",
      TRUE ~ "Tie"  # In case of a tie or missing data
    )
  )

# unique(et$Resp)
# unique(et$Fit_Method)
# unique(et$condit)
# unique(et$x)

@@

```{r, fig.width=12, fig.height=11}
source(here::here("Functions/fun_indv_fit.R"))

plot_sampled_posterior(ind_fits)
plot_indv_posterior(ind_fits_df)
group_predictive_plots(ind_fits_df)
group_best_plots(ind_fits_df)
indv_best_plots(ind_fits_df)
indv_predictive_plots(ind_fits_df, ids2)
indv_predictive_plots(ind_fits_df, idsBad)
```

pdf(here::here("assets/tmp_plots/test.pdf"))
plot_sampled_posterior(ind_fits)
plot_indv_posterior(ind_fits_df)
group_predictive_plots(ind_fits_df)
group_best_plots(ind_fits_df)
indv_best_plots(ind_fits_df)
indv_predictive_plots(ind_fits_df, ids2)
indv_predictive_plots(ind_fits_df, idsBad)
dev.off()

```{r fig.width=10, fig.height=8}

post_dat_l |> group_by(id,condit,Fit_Method, Resp,x) |>
  summarise(val=mean(val)) |>
  ggplot(aes(x = x, y = val, fill=Resp)) + 
  stat_bar + 
  facet_wrap(condit~Fit_Method, scales="free") + 
  labs(title = "Full Posterior")

post_dat_l |> group_by(id,condit,Fit_Method, Resp,x) |>
  filter(rank==1) |>
  summarise(val=mean(val)) |>
  ggplot(aes(x = x, y = val, fill=Resp)) + 
  stat_bar + 
  facet_wrap(condit~Fit_Method, scales="free") + 
  labs(title = "Full Posterior")
```

```{r fig.width=11, fig.height=15}

post_dat_l |> filter(id %in% ids2, rank==1) |> ggplot(aes(x = x, y = val, fill=Resp)) + 
  stat_bar + 
  facet_wrap(Fit_Method~id, scales="free") + 
  labs(title = "Single Best Fit")

post_dat_l |> filter(id %in% ids2) |> ggplot(aes(x = x, y = val, fill=Resp)) + 
  stat_bar + 
  facet_wrap(Fit_Method~id, scales="free") + 
  labs(title = "Average of posterior predictive distributions")

post_dat_l |> filter(id %in% ids2, rank<=10) |> ggplot(aes(x = x, y = val, fill=Resp)) + 
  stat_bar + 
  facet_wrap(Fit_Method~id, scales="free") + 
  labs(title = "10 best fits")

@@

r_all <- post_dat_l |> filter(id %in% ids) |> mutate(dist="All")
r1 <- post_dat_l |> filter(id %in% ids, rank==1) |> mutate(dist="best")
r_10 <- post_dat_l |> filter(id %in% ids, rank<=10) |> mutate(dist="top10")

d <- bind_rows(r_all,r1)

d |> ggplot(aes(x = x, y = val, fill=Resp)) + 
  stat_bar + 
  ggh4x::facet_nested_wrap(id~Fit_Method~dist, scales="free",ncol=6) + 
  labs(title = "Average of posterior predictive distributions")

library(dtplyr)

post_dt <- ind_fits_df |> select(sim_dat,rank) |> unnest(sim_dat) # > dtplyr::lazy_dt()
post_dat <- ind_fits_df |> select(sim_dat,rank) |> unnest(sim_dat) |> as.data.table()

post_dat1 <- filter(post_dat,expMode2=="Test") |>
  group_by(id,condit,x) |>
  mutate(avg_y=mean(y)) |> 
  #relocate(id,condit,Model,Fit_Method,expMode2,tr,rank,c,lr,x,y,avg_y) |>
  show_query()

t1 <- system.time({ 
post_dat1 <- filter(post_dat,expMode2=="Test") |>
  group_by(id,condit,x) |>
  mutate(avg_y=mean(y)) |> 
  relocate(id,condit,Model,Fit_Method,expMode2,tr,rank,c,lr,x,y,avg_y) |>
  pivot_longer(
    cols = c(pred, y),
    names_to = "Resp",
    values_to = "val"
  ) %>%
  mutate(
    Resp = case_when(
      Resp == "y" ~ "Observed",
      Model == "ALM" ~ "ALM",
      Model == "EXAM" ~ "EXAM"
    )
  ) |> arrange(tr,id,Resp) 
})

t2 <- system.time({ 
post_dat2 <- setorder(melt(setcolorder(post_dat[expMode2 == "Test"][, `:=`(avg_y = mean(y)), 
    by = .(id, condit, x)], c("id", "condit", "Model", "Fit_Method", 
"expMode2", "tr", "rank", "c", "lr", "x", "y", "avg_y", "pred", 
"resid")), measure.vars = c("pred", "y"), variable.name = "Resp", 
    value.name = "val", variable.factor = FALSE)[, `:=`(Resp = fcase(Resp == 
    "y", "Observed", Model == "ALM", "ALM", Model == "EXAM", 
    "EXAM")), by = .(id, condit, x)], tr, id, Resp, na.last = TRUE)
})

show_query()
post_dat_l <- post_dat %>%
